PaperMelanoma

### 2022-06-09 Code/PaperMelanoma.Rmd
Added tests

### Recode
Set Rdamcrit of negative SN patients equal to 0
```{r, eval=TRUE}
sel.data$Rdamcrit[sel.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

@@

Load imputed data set
```{r, eval=TRUE}
load('Z:/Project Melanoom/PaperMelanoma/Data/imputed.data.RData')
complete.imputed.data <- mice::complete(imputed.data, m)
summary(complete.imputed.data)
missings.table(complete.imputed.data)
```

# Set as survival data
```{r, eval=TRUE}
S.Rec <- survival::Surv(exp(complete.imputed.data$logRec.time), sel.data$Recurrence=="recurred")
S.MSM <- survival::Surv(exp(complete.imputed.data$logFU.time), sel.data$MSM)
```

@@

# Median follow-up time
```{r, eval=TRUE}
median.FU.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~1, data=complete.imputed.data, reverse=TRUE))
tab.FU.time <- c("total", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                 paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                        sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
median.FU.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~Center, data=complete.imputed.data, reverse=TRUE))
for (center in sort(as.character(unique(sel.data$Center)))){
  med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
  med.FU <- med.FU.table[3,"quantile"]
  lower.FU <- med.FU.table[3,"lower"]
  upper.FU <- med.FU.table[3,"upper"]
  tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
}
write.table(as.data.frame(tab.FU.time),file='Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt',
            row.names=FALSE)
tab.FU.time
```

@@

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
c.Breslow <- mean(log(complete.imputed.data$Breslow))
form.BS.Rec <- S.Rec ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec <- Hmisc::fit.mult.impute(form.BS.Rec,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                           pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
round(stats::anova(f.mi.BS.Rec)[order(stats::anova(f.mi.BS.Rec)[, 'P']),], 3)
```

@@

# Refitted final model for melanoma specific mortality
```{r, eval=TRUE}
form.BS.MSM.refit <- S.MSM ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.MSM.refit <- Hmisc::fit.mult.impute(form.BS.MSM.refit,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                                  pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
round(anova(f.mi.BS.MSM.refit)[order(anova(f.mi.BS.MSM.refit)[, 'P']),], 3)
f.mi.BS.MSM.refit
summary(f.mi.BS.MSM.refit)
stats::anova(f.mi.BS.MSM.refit)
```

@@

##### Final model
```{r, eval=TRUE}
for (model in c('Rec', 'MSM', 'MSM.refit')){
  if (model=="MSM"){
    mult.factor <- MSM.cal.fact # if MSM, multiply by MSM.cal.fact
    f.mi.BS <- eval(parse(text=paste0('f.mi.BS.Rec'))) # use recurrence model if calibrated MSM
  }
  else{
    mult.factor <- 1 # if Rec or MSM.refit, don't multiply
    f.mi.BS <- eval(parse(text=paste0('f.mi.BS.', model)))
  }

  CI.lower.neg <- exp(mult.factor*confint(f.mi.BS)[,1])
  CI.upper.neg <- exp(mult.factor*confint(f.mi.BS)[,2])
  CI.neg <- paste0("[", sprintf("%.2f", CI.lower.neg), "; ", sprintf("%.2f", CI.upper.neg), "]")

@@

  CI.pos <- paste0("[", sprintf("%.2f", exp(mult.factor*confint(f.mi.BS)[7,1])*exp(mult.factor*confint(f.mi.BS)[9,1])),
                   "; ", sprintf("%.2f", exp(mult.factor*confint(f.mi.BS)[7,2])*exp(mult.factor*confint(f.mi.BS)[9,2])), "]")

  BS.table <- cbind(c(NA, sprintf("%.2f", exp(mult.factor*f.mi.BS$coefficients[2:7])), NA), # COEF NEG
                    c(NA, CI.neg[2:7], NA), # CI NEG
                    sprintf("%.2f", exp(mult.factor*f.mi.BS$coefficients[1:8])*c(rep(1, 6), exp(mult.factor*f.mi.BS$coefficients[9]), 1)), #COEF POS
                    c(CI.neg[1:6], CI.pos, CI.neg[8]), # CI POS
                    c(NA, sprintf("%.1f", anova.BS[3:5]), rep(NA, 2), sprintf("%.1f", anova.BS[c(6, 8)])), # CHI2 NEG
                    c(sprintf("%.1f", anova.BS[1]), rep(NA, 5), # CHI2 POS
                      sprintf("%.1f", anova.BS[9]), NA))

@@

# Plot calibration of all centers combined and all centers seperate for recurrence and melanoma specific mortality refit and write results to Excel
```{r, eval=TRUE}
horizon <- 5
for (model in list('Rec', 'MSM.refit')){
  if (model == "Rec"){
    # plot predict
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Rec.png"))
    plot.predict <- plot(Predict(f.mi.full.Rec))
    show(plot.predict)
    grDevices::dev.off()

@@

# replace NA by blank
final.full.table.Rec[is.na(final.full.table.Rec)] <- ""
final.full.table.MSM.refit[is.na(final.full.table.MSM.refit)] <- ""
final.BS.table.Rec[is.na(final.BS.table.Rec)] <- ""
final.BS.table.MSM[is.na(final.BS.table.MSM)] <- ""
final.BS.table.MSM["C-index MSM",] <- final.BS.table.Rec[nrow(final.BS.table.Rec),]
final.BS.table.MSM.refit[is.na(final.BS.table.MSM.refit)] <- ""

# save tables
write.table(final.full.table.Rec,file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.Rec.txt')
write.table(final.full.table.MSM.refit,file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.MSM.refit.txt')
write.table(final.BS.table.Rec,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.Rec.txt')
write.table(final.BS.table.MSM,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.txt')
write.table(final.BS.table.MSM.refit,file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.refit.txt')
final.full.table.Rec
final.full.table.MSM.refit
final.BS.table.Rec
final.BS.table.MSM
final.BS.table.MSM.refit
```

@@

# Baseline hazard
rc.Rec <- (1/attributes(nom.0)$info$sc)
int.Rec <- attributes(nom.0)$info$Intercept
f.Rec.basehaz <- basehaz(f.mi.BS.Rec,TRUE)
h0.Rec <- f.Rec.basehaz$hazard[f.Rec.basehaz$time==max(f.Rec.basehaz$time[f.Rec.basehaz$time<=horizon])]
lp.range <- int.Rec+rc.Rec*(0:10)*2
par(mar = rep(0, 4))
nom <- nomogram(f.mi.BS.Rec, maxscale=10,
              Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class,
              lp=TRUE, fun.at=fun.event(lp=lp.range, h0=h0.Rec))
attributes(nom)$names[attributes(nom)$names=="Age.SN"]<-"Age"
attributes(nom)$names[attributes(nom)$names=="Loc_CAT"]<-"Location"
attributes(nom)$names[attributes(nom)$names=="Rdamcrit"]<-"Tumor burden"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Positive)"]<-"Breslow + positive SN"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Negative)"]<-"Breslow + negative SN"
plot(nom, total.sep.page=TRUE, col.grid=gray(c(0.8, 0.95)))
nom.0
nom

# Slope and intercept of recurrence and MSM
int.Rec
# check that int is same as this:
log(20)*f.mi.BS.Rec$coef["Age.SN"]+log(0.1)*f.mi.BS.Rec$coef["Breslow"]+log(0.01)*f.mi.BS.Rec$coef["Rdamcrit"]-f.mi.BS.Rec$center
```

@@

Breslow.trunc <- complete.imputed.data$Breslow
Breslow.trunc[complete.imputed.data$Breslow<0.1] <- .1
Breslow.trunc[complete.imputed.data$Breslow>7] <- 7
lp.score.Rec <- predict(f.mi.BS.Rec,newdata=data.frame(SNstatus=complete.imputed.data$SNstatus,
                                                Age.SN=complete.imputed.data$Age.SN,
                                                Ulceration=complete.imputed.data$Ulceration,
                                                Loc_CAT=complete.imputed.data$Loc_CAT,
                                                Breslow=Breslow.trunc,
                                                Rdamcrit=complete.imputed.data$Rdamcrit))
mean(lp.score.Rec)
score.Rec <- round((lp.score.Rec-int.Rec)/rc.Rec,0)
max.breaks <- max.range+1
h.Rec <- hist(score.Rec,plot=FALSE,breaks=0:max.breaks,right=FALSE)

@@

# Check mean event rates of risk distribution
```{r, eval=TRUE}
# Recurrence
sum(h.Rec$density*fun.event(lp=lp.points.Rec, h0=h0.Rec))
# equal to the 5-year survival probability, kaplanmeier op 5 jaar, survfit op 5 jaar
SF<-survfit(S.Rec~1,data=sel.data)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(fun.event(lp=lp.score.Rec, h0=h0.Rec))

@@

# Plot of risk score distribution
```{r, eval=TRUE, message=FALSE}
h.plot <- h.Rec
h.plot$density <- 100*h.Rec$density
p.Rec <- fun.event(lp=lp.points.Rec, h0=h0.Rec)
p.MSM <- fun.event(lp=lp.points.MSM, h0=h0.MSM)
x.lim <- c(0, max.range+1)
y.lim <- c(0, 100)
y.lim.hist <- c(0, 25)
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Risk.distribution.png"))
par(mar = c(5,5,2,5))
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col='white')
abline(h=(0:30)*1,col="light grey",lty=2)
axis(side = 4)
mtext(side = 4, line = 3, 'Risk score distribution (%)')
par(new = TRUE)
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")
par(new = TRUE)
plot(points.range,p.Rec*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=15, xlab="Risk score",ylab="Predicted 5-year risk (%)")
points(points.range,p.MSM*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=16)
legend(x=0,y=100,legend=c("Recurrence","Melanoma-specific mortality"),pch=c(15,16),cex=.9,bg='white')
grDevices::dev.off()
# table of risk distribution
cbind(points.range, round(p.Rec*100, 1), round(p.MSM*100, 1))[c(1, 9, 10, 12, 13),]

# save for Rshiny app
coef.Rec <- coef(f.mi.BS.Rec)
center.Rec <- f.mi.BS.Rec$center
save(h0.Rec, h0.MSM, coef.Rec, c.Breslow, center.Rec,
     file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'), compress=TRUE)
load(file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'))
```

# Example to check probabilities with app
```{r, eval=TRUE}
patient.nr <- 115 # 395 and 396 = Negative, 397 and 398 = Positive, 30 = Head & neck, 115 has large Breslow
data.to.be.imputed[patient.nr, c("SNstatus", "Age.SN", "Ulceration", "Loc_CAT", "Breslow", "Rdamcrit")]
lp.score.Rec.true <- predict(f.mi.BS.Rec)[patient.nr]
lp.score.Rec.try <- predict(f.mi.BS.Rec, newdata=data.frame(SNstatus=data.to.be.imputed[patient.nr,"SNstatus"],
                                                            Age.SN=data.to.be.imputed[patient.nr,"Age.SN"],
                                                            Ulceration=data.to.be.imputed[patient.nr,"Ulceration"],
                                                            Loc_CAT=data.to.be.imputed[patient.nr,"Loc_CAT"],
                                                            Breslow=data.to.be.imputed[patient.nr,"Breslow"],
                                                            Rdamcrit=data.to.be.imputed[patient.nr,"Rdamcrit"]),
                                                            type="lp")
p.Rec.example <- fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
p.MSM.example <- fun.event(lp=lp.score.Rec.try, h0=h0.MSM)

@@

# manual calculation
coef.Rec <- coef(f.mi.BS.Rec)
lp.manual <- coef.Rec["SNstatus=Positive"]*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)+
              coef.Rec["Age.SN"]*log(data.to.be.imputed[patient.nr,"Age.SN"])+
              coef.Rec["Ulceration=Yes"]*(as.numeric(data.to.be.imputed[patient.nr,"Ulceration"])-1)+
              coef.Rec["Loc_CAT=leg"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="leg")+
              coef.Rec["Loc_CAT=trunk"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="trunk")+
              coef.Rec["Loc_CAT=head & neck"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="head & neck")+
              coef.Rec["Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)+
              coef.Rec["Rdamcrit"]*log(data.to.be.imputed[patient.nr,"Rdamcrit"])+
              coef.Rec["SNstatus=Positive * Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)
center.Rec <- f.mi.BS.Rec$center

>>>>

### Recode
Set Rdamcrit of negative SN patients equal to 0. 11 patients with positive SN have SN tumour burden of exactly 1
```{r, eval=TRUE}
sum(sel.data$Rdamcrit==1&sel.data$SNstatus=="Positive"&!is.na(sel.data$Rdamcrit))
sel.data$Rdamcrit[sel.data$SNstatus=="Negative"] <- 1 # since log(1) = 0
```

@@

Load imputed data set
```{r, eval=TRUE}
load('Z:/Project Melanoom/PaperMelanoma/Data/imputed.data.RData')
last.imputed.data.set <- mice::complete(imputed.data, m)
summary(last.imputed.data.set)
missings.table(last.imputed.data.set)
```

# Set as survival data
```{r, eval=TRUE}
S.Rec <- survival::Surv(exp(last.imputed.data.set$logRec.time), sel.data$Recurrence=="recurred")
S.MSM <- survival::Surv(exp(last.imputed.data.set$logFU.time), sel.data$MSM)
```

@@

# Median follow-up time
```{r, eval=TRUE}
median.FU.time.total <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~1, data=last.imputed.data.set, reverse=TRUE))
tab.FU.time <- c("total", paste(sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "quantile"]),
                 paste0("[", sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "lower"]), "; ",
                        sprintf("%.2f", median.FU.time.total$quantiles.survival[3, "upper"]), "]")))
median.FU.time <- stats::quantile(prodlim::prodlim(prodlim::Hist(exp(logFU.time), MSM)~Center, data=last.imputed.data.set, reverse=TRUE))
for (center in sort(as.character(unique(sel.data$Center)))){
  med.FU.table <- eval(parse(text=paste0("median.FU.time$`Center=", center, "`")))
  med.FU <- med.FU.table[3,"quantile"]
  lower.FU <- med.FU.table[3,"lower"]
  upper.FU <- med.FU.table[3,"upper"]
  tab.FU.time <- rbind(tab.FU.time, c(center, paste(sprintf("%.2f", med.FU), paste0("[", sprintf("%.2f", lower.FU), "; ", sprintf("%.2f", upper.FU), "]"))))
}
write.table(as.data.frame(tab.FU.time),file='Z:/Project Melanoom/PaperMelanoma/Results/med.FU.table.txt',
            row.names=FALSE)
tab.FU.time
```

@@

# Final model for recurrence selected by backward selection with p = 0.01
```{r, eval=TRUE}
c.Breslow <- mean(log(last.imputed.data.set$Breslow))
form.BS.Rec <- S.Rec ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.Rec <- Hmisc::fit.mult.impute(form.BS.Rec,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                           pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
# round(stats::anova(f.mi.BS.Rec)[order(stats::anova(f.mi.BS.Rec)[, 'P']),], 3)
stats::anova(f.mi.BS.Rec)
```

@@

# Refitted final model for melanoma specific mortality
```{r, eval=TRUE}
form.BS.MSM.refit <- S.MSM ~
  SNstatus +
  log(Age.SN) +
  Ulceration +
  Loc_CAT +
  I(log(Breslow)-c.Breslow) +
  log(Rdamcrit) +
  SNstatus:I(log(Breslow)-c.Breslow)
f.mi.BS.MSM.refit <- Hmisc::fit.mult.impute(form.BS.MSM.refit,cph,xtrans=imputed.data,data=data.to.be.imputed,n.impute=m,
                                  pr=FALSE,fit.reps=TRUE,y=TRUE,x=TRUE,se.fit=TRUE)
# round(anova(f.mi.BS.MSM.refit)[order(anova(f.mi.BS.MSM.refit)[, 'P']),], 3)
f.mi.BS.MSM.refit
summary(f.mi.BS.MSM.refit)
stats::anova(f.mi.BS.MSM.refit)
```

@@

##### Final model
```{r, eval=TRUE}
for (model in c('Rec', 'MSM.refit')){
  f.mi.BS <- eval(parse(text=paste0('f.mi.BS.', model)))

  CI.lower.neg <- exp(confint(f.mi.BS)[,1])
  CI.upper.neg <- exp(confint(f.mi.BS)[,2])
  CI.neg <- paste0("[", sprintf("%.2f", CI.lower.neg), "; ", sprintf("%.2f", CI.upper.neg), "]")

@@

  CI.pos <- paste0("[", sprintf("%.2f", exp(confint(f.mi.BS)[7,1])*exp(confint(f.mi.BS)[9,1])),
                   "; ", sprintf("%.2f", exp(confint(f.mi.BS)[7,2])*exp(confint(f.mi.BS)[9,2])), "]")

  BS.table <- cbind(c(NA, sprintf("%.2f", exp(f.mi.BS$coefficients[2:7])), NA), # COEF NEG
                    c(NA, CI.neg[2:7], NA), # CI NEG
                    sprintf("%.2f", exp(f.mi.BS$coefficients[1:8])*c(rep(1, 6), exp(f.mi.BS$coefficients[9]), 1)), #COEF POS
                    c(CI.neg[1:6], CI.pos, CI.neg[8]), # CI POS
                    c(NA, sprintf("%.1f", anova.BS[3:5]), rep(NA, 2), sprintf("%.1f", anova.BS[c(6, 8)])), # CHI2 NEG
                    c(sprintf("%.1f", anova.BS[1]), rep(NA, 5), # CHI2 POS
                      sprintf("%.1f", anova.BS[9]), NA))

@@

# Plot calibration of all centers combined and all centers seperate 
for recurrence and melanoma specific mortality refit 
write results to Excel
```{r, eval=TRUE}
horizon <- 5
for (model in list('Rec', 'MSM.refit')){
  if (model == "Rec"){
    # plot predict
    png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Figures/plot.predict.Rec.png"))
    plot.predict <- plot(Predict(f.mi.full.Rec))
    show(plot.predict)
    grDevices::dev.off()

@@

# replace NA by blank
final.full.table.Rec[is.na(final.full.table.Rec)] <- ""
final.full.table.MSM.refit[is.na(final.full.table.MSM.refit)] <- ""
final.BS.table.Rec[is.na(final.BS.table.Rec)] <- ""
final.BS.table.MSM.refit[is.na(final.BS.table.MSM.refit)] <- ""

# set rownames
final.full.table.Rec <- as.data.frame(final.full.table.Rec)
row.names(final.full.table.Rec) <- c("Positive SN status", "Male",
                                    "Age", "Ulceration", "Location",
                                    "Arm", "Leg", "Trunk", "Head and neck",
                                    "Histology", "SSM", "NM", "ALM", "Other",
                                    "Breslow", "Multiple fields", 
                                    "Total number of nodes",
                                    "SN tumour burden", 
                                    "Location metastasis in lymph node",
                                    "Subcap", "Combined", "Parenchymal",
                                    "Multifocal", "Extensive",
                                    rownames(final.full.table.Rec)[25:26])
final.full.table.MSM.refit <- as.data.frame(final.full.table.MSM.refit)
row.names(final.full.table.MSM.refit) <- c("Positive SN status", "Male",
                                          "Age", "Ulceration", "Location",
                                          "Arm", "Leg", "Trunk", "Head and neck",
                                          "Histology", "SSM", "NM", "ALM", "Other",
                                          "Breslow", "Multiple fields", 
                                          "Total number of nodes",
                                          "SN tumour burden", 
                                          "Location metastasis in lymph node",
                                          "Subcap", "Combined", "Parenchymal",
                                          "Multifocal", "Extensive",
                                          row.names(final.full.table.Rec)[25])
final.BS.table.Rec <- as.data.frame(final.BS.table.Rec)
row.names(final.BS.table.Rec) <- c("Positive SN status", "Age", "Ulceration",
                                   "Location", "Arm", "Leg", "Trunk", 
                                   "Head and neck",
                                   "Breslow", "SN tumour burden", 
                                   row.names(final.BS.table.Rec)[11:12])
final.BS.table.MSM.refit <- as.data.frame(final.BS.table.MSM.refit)
row.names(final.BS.table.MSM.refit) <- c("Positive SN status", "Age", "Ulceration",
                                   "Location", "Arm", "Leg", "Trunk", 
                                   "Head and neck",
                                   "Breslow", "SN tumour burden", 
                                   row.names(final.BS.table.MSM.refit)[11])

# save tables
write.table(final.full.table.Rec,
            file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.Rec.txt', sep=",")
write.table(final.full.table.MSM.refit,
            file='Z:/Project Melanoom/PaperMelanoma/Results/final.full.table.MSM.refit.txt', sep=",")
write.table(final.BS.table.Rec,
            file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.Rec.txt', sep=",")
write.table(final.BS.table.MSM.refit,
            file='Z:/Project Melanoom/PaperMelanoma/Results/final.BS.table.MSM.refit.txt', sep=",")

# show tables
final.full.table.Rec
final.full.table.MSM.refit
final.BS.table.Rec
final.BS.table.MSM.refit
```

@@

# Baseline hazard
rc.Rec <- (1/attributes(nom.0)$info$sc)
int.Rec <- attributes(nom.0)$info$Intercept
f.Rec.basehaz <- basehaz(f.mi.BS.Rec,TRUE)
h0.Rec <- f.Rec.basehaz$hazard[f.Rec.basehaz$time==max(f.Rec.basehaz$time[f.Rec.basehaz$time<=horizon])]
lp.range <- int.Rec+rc.Rec*(0:10)*2
par(mar = rep(0, 4))
nom <- nomogram(f.mi.BS.Rec, maxscale=10,
              Breslow=Breslow.class, Rdamcrit=Rdamcrit.class, Age.SN=Age.class,
              lp=TRUE, fun.at=fun.event(lp=lp.range, h0=h0.Rec))
attributes(nom)$names[attributes(nom)$names=="Age.SN"]<-"Age"
attributes(nom)$names[attributes(nom)$names=="Loc_CAT"]<-"Location"
attributes(nom)$names[attributes(nom)$names=="Rdamcrit"]<-"Tumor burden"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Positive)"]<-"Breslow + positive SN"
attributes(nom)$names[attributes(nom)$names=="Breslow (SNstatus=Negative)"]<-"Breslow + negative SN"
plot(nom, total.sep.page=TRUE, col.grid=gray(c(0.8, 0.95)))
nom.0
nom
```

@@

Breslow.trunc <- last.imputed.data.set$Breslow
Breslow.trunc[last.imputed.data.set$Breslow<0.1] <- .1
Breslow.trunc[last.imputed.data.set$Breslow>7] <- 7
lp.score.Rec <- predict(f.mi.BS.Rec,newdata=data.frame(SNstatus=last.imputed.data.set$SNstatus,
                                                Age.SN=last.imputed.data.set$Age.SN,
                                                Ulceration=last.imputed.data.set$Ulceration,
                                                Loc_CAT=last.imputed.data.set$Loc_CAT,
                                                Breslow=Breslow.trunc,
                                                Rdamcrit=last.imputed.data.set$Rdamcrit))
mean(lp.score.Rec)
score.Rec <- round((lp.score.Rec-int.Rec)/rc.Rec,0)
max.breaks <- max.range+1
h.Rec <- hist(score.Rec,plot=FALSE,breaks=0:max.breaks,right=FALSE)

@@

# Plot of risk score distribution
```{r, eval=TRUE, message=FALSE}
h.plot <- h.Rec
h.plot$density <- 100*h.Rec$density
p.Rec <- fun.event(lp=lp.points.Rec, h0=h0.Rec)
p.MSM <- fun.event(lp=lp.points.MSM, h0=h0.MSM)
x.lim <- c(0, max.range+1)
y.lim <- c(0, 100)
y.lim.hist <- c(0, 25)
png(file=paste0("Z:/Project Melanoom/PaperMelanoma/Results/Risk.distribution.png"))
par(mar = c(5,5,2,5))
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col='white')
abline(h=(0:30)*1,col="light grey",lty=2)
axis(side = 4)
mtext(side = 4, line = 3, 'Risk score distribution (%)')
par(new = TRUE)
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")
par(new = TRUE)
plot(points.range,p.Rec*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=15, xlab="Risk score",ylab="Predicted 5-year risk (%)")
points(points.range,p.MSM*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=16)
legend(x=0,y=100,legend=c("Recurrence","Melanoma-specific mortality"),pch=c(15,16),cex=.9,bg='white')
grDevices::dev.off()
```

```{r, eval=TRUE, echo=FALSE}
par(mar = c(5,5,2,5))
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col='white')
abline(h=(0:30)*1,col="light grey",lty=2)
axis(side = 4)
mtext(side = 4, line = 3, 'Risk score distribution (%)')
par(new = TRUE)
plot(h.plot,freq=FALSE,axes=FALSE, xlab=NA, xlim=x.lim, ylab=NA,ylim=y.lim.hist,main=NA,col="white")

par(new = TRUE)
plot(points.range,p.Rec*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=15, xlab="Risk score",ylab="Predicted 5-year risk (%)")
points(points.range,p.MSM*100,type="b",lwd=1,xlim=x.lim,ylim=y.lim,pch=16)
legend(x=0,y=100,legend=c("Recurrence","Melanoma-specific mortality"),pch=c(15,16),cex=.9,bg='white')
```

```{r, eval=TRUE}
# reset margins
par(mar = c(5,5,2,5))

# table of risk distribution
cbind(points.range, round(p.Rec*100, 1), round(p.MSM*100, 1))[c(1, 9, 10, 12, 13),]

# save for Rshiny app
coef.Rec <- stats::coef(f.mi.BS.Rec)
center.Rec <- f.mi.BS.Rec$center
save(h0.Rec, h0.MSM, coef.Rec, c.Breslow, center.Rec,
     file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'), compress=TRUE)
load(file=paste0('Z:/Project Melanoom/PaperMelanoma/Results/h0.Rec.MSM.Rdata'))
```

# Test if model is correct
### Check calibration
```{r, eval=TRUE}
# Slope and intercept of recurrence and MSM
int.Rec
# check that int is same as this:
log(20)*f.mi.BS.Rec$coef["Age.SN"]+log(0.1)*f.mi.BS.Rec$coef["Breslow"]+log(0.01)*f.mi.BS.Rec$coef["Rdamcrit"]-f.mi.BS.Rec$center
```

### Check mean event rates of risk distribution
```{r, eval=TRUE}
# Recurrence
sum(h.Rec$density*fun.event(lp=lp.points.Rec, h0=h0.Rec))
# equal to the 5-year survival probability, kaplanmeier op 5 jaar, survfit op 5 jaar
SF<-survfit(S.Rec~1,data=sel.data)
100*(1-SF$surv[SF$time==max(SF$time[SF$time<=horizon])])
mean(fun.event(lp=lp.score.Rec, h0=h0.Rec))

@@

### Example to check probabilities with app
```{r, eval=TRUE}
patient.nr <- 115 # 395 and 396 = Negative, 397 and 398 = Positive, 30 = Head & neck, 115 has large Breslow
data.to.be.imputed[patient.nr, c("SNstatus", "Age.SN", "Ulceration", "Loc_CAT", "Breslow", "Rdamcrit")]
lp.score.Rec.true <- predict(f.mi.BS.Rec)[patient.nr]
lp.score.Rec.try <- predict(f.mi.BS.Rec, newdata=data.frame(SNstatus=data.to.be.imputed[patient.nr,"SNstatus"],
                                                            Age.SN=data.to.be.imputed[patient.nr,"Age.SN"],
                                                            Ulceration=data.to.be.imputed[patient.nr,"Ulceration"],
                                                            Loc_CAT=data.to.be.imputed[patient.nr,"Loc_CAT"],
                                                            Breslow=data.to.be.imputed[patient.nr,"Breslow"],
                                                            Rdamcrit=data.to.be.imputed[patient.nr,"Rdamcrit"]),
                                                            type="lp")
p.Rec.example <- fun.event(lp=lp.score.Rec.try, h0=h0.Rec)
p.MSM.example <- fun.event(lp=lp.score.Rec.try, h0=h0.MSM)

@@

# manual calculation
coef.Rec <- stats::coef(f.mi.BS.Rec)
lp.manual <- coef.Rec["SNstatus=Positive"]*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)+
              coef.Rec["Age.SN"]*log(data.to.be.imputed[patient.nr,"Age.SN"])+
              coef.Rec["Ulceration=Yes"]*(as.numeric(data.to.be.imputed[patient.nr,"Ulceration"])-1)+
              coef.Rec["Loc_CAT=leg"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="leg")+
              coef.Rec["Loc_CAT=trunk"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="trunk")+
              coef.Rec["Loc_CAT=head & neck"]*as.numeric(data.to.be.imputed[patient.nr,"Loc_CAT"]=="head & neck")+
              coef.Rec["Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)+
              coef.Rec["Rdamcrit"]*log(data.to.be.imputed[patient.nr,"Rdamcrit"])+
              coef.Rec["SNstatus=Positive * Breslow"]*(log(data.to.be.imputed[patient.nr,"Breslow"])-c.Breslow)*(as.numeric(data.to.be.imputed[patient.nr,"SNstatus"])-1)
center.Rec <- f.mi.BS.Rec$center
